---
title: "Homework 8"
author: "Meiruo Xiang"
date: "10/29/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dpi=300,fig.width=7)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
```


# 5.3.3 Orstein-Uhlenbeck Process

### 1. Orstein-Uhlenbeck Process proof

Orstein-Uhlenbeck Process: $dr(t)=\alpha(b-r(t))dt+\sigma dW_t$

let $f(x_t, t)=r(t)e^{\alpha t}$

By, Ito's fomula,

$$
\begin{aligned}
df(r(t), t)&=\frac{\partial}{\partial t}(r(t)e^{\alpha t})dt+\frac{\partial}{\partial r(t)}(r(t)e^{\alpha t})dr(t)\\
&=r(t)e^{\alpha t}\alpha dt+e^{\alpha t}dr(t)\\
&=r(t)e^{\alpha t}\alpha dt+e^{\alpha t}*[\alpha(b-r(t))dt+\sigma dW_t]\\
&=\alpha b e^{\alpha t} dt + e^{\alpha t}\sigma dW_t
\end{aligned}
$$

Integral form 0 to t,

$$
\begin{aligned}
r(t)e^{\alpha t}&=r(0)+\int_0^t \alpha b e^{\alpha s} ds+ \int_0^t e^{\alpha s}\sigma dW_s\\
&=r(0)+ b(e^{\alpha t}-1)+\sigma \int_0^t e^{\alpha s} dW_s
\end{aligned}
$$

so 

$$
\begin{aligned}
r(t)&=r(0)e^{-\alpha t}+ b(1-e^{-\alpha t})+\sigma \int_0^t e^{-\alpha (t-s)} dW_s\\
r(t+\Delta)&=r(0)e^{-\alpha (t+\Delta)}+ b(1-e^{-\alpha (t+\Delta)})+\sigma \int_0^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s\\
r(t+\Delta)-e^{-\alpha \Delta}r(t)&=b(1-e^{-\alpha \Delta})+\sigma \int_t^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s\\
\end{aligned}
$$
By property of Brownian motion,
$W(s+\Delta s)-W(s)$ follows a normal distribution with mean 0.
so $E[\int_t^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s]=0$

By Ito's isometry,

$$
\begin{aligned}
Var[\int_t^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s]
&=E[(\int_t^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s)^2]\\
&=E[(\int_0^{\Delta} e^{-\alpha (\Delta-y)} dW_y)^2]\\
&=E[\int_0^{\Delta} e^{-2\alpha (\Delta-y)} dy]\\
&=e^{-2\alpha \Delta} \frac{1}{2\alpha}(e^{2\alpha \Delta}-1)\\
&=\frac{1}{2\alpha}(1-e^{-2\alpha \Delta})\\
\end{aligned}
$$

Therefore,
$$
\begin{aligned}
r(t+\Delta)&=e^{-\alpha \Delta}r(t)+b(1-e^{-\alpha \Delta})+\sigma \int_t^{t+\Delta} e^{-\alpha (t+\Delta-s)} dW_s\\
&=e^{-\alpha \Delta}r(t)+b(1-e^{-\alpha \Delta})+\sigma  \sqrt{\frac{1}{2\alpha}(1-e^{-2\alpha \Delta})}*Z, Z\sim N(0,1)
\end{aligned}
$$



### 2. Implement a random walk.

```{r}
orstein <- function(alpha, sigma, b, r0 = 1, tmax = 500, delta = 1/500) {   
  n <- tmax / delta
  r <- numeric(n)
  r[1] <- r0
  for(i in 1:n) {   
    r[i+1] <- exp(-alpha * delta) * r[i] + 
              b * (1 - exp(-alpha * delta)) + 
              sigma / sqrt(2 * alpha) * 
              sqrt((1 - exp(-2 * alpha * delta))) * rnorm(1)
  }
  return(r[-1])
}

alpha <- c(0.1, 1, 5)
sigma <- c(0.1, 0.2, 0.5)
b <- c(-5, 5)

# plot for each combination of parameters
par(mfrow=c(3,3))
l <- 1
for (i in 1:2) {
  for (j in 1:3) {
    for (k in 1:3) {
      title <- paste('alpha=', alpha[j],
                     'sigma=', sigma[k],
                     'b=', b[i])
      plot(ts(orstein(alpha[j], sigma[k], b[i]), frequency=500),type='l', 
            ylab='Sample path', main = title, col = l)
      l <- l + 1
    }    
  }  
}

```
 

As $\alpha$ increase, the variance of path gets smaller, the line looks flatter.
As $\sigma$ increase, the fluctuation of the oscillation gets larger, the amplitude is larger.

### 3. Use the Euler–Maruyama method to approximate a simulation from the process.
```{r}
euler <- function(alpha = 0.1, sigma = 0.1, b = 5, r0, tmax = 1, delta) { 
  size <- 1000
  n <- tmax/delta
  r <- matrix(0, size, n + 1)
  r[, 1] <- r0
  for (i in 1:size) {   
      for (j in 1:n) {  
        mean <- r[i, j] + alpha * (b - r[i, j]) * delta
        sd <- sigma * delta
        r[i, j + 1] <- rnorm(1, mean = mean, sd = sd)  
      }
  }
  return(r[, dim(r)[2]])
}


r0 <- 1
delta <- c(1, 0.5, 0.1, 0.01)
alpha <- 0.1
sigma <- 0.1
b <- 5
mu <- exp(-alpha * delta[1]) * r0 + b * (1 - exp(-alpha * delta[1]))
sd <- sigma * sqrt((1 - exp(-2 * alpha * delta[1])) / (2 * alpha))


for(i in 1:length(delta)){
  simulate <- euler(r0 = r0, delta = delta[i])
  title <- paste("Euler approximation with delta = ", delta[i])
  hist(simulate, freq = F,
       xlab = 'r', main = title)
  lines(density(simulate), col = 1, lwd = 1)
  curve(dnorm(x, mean=mu, sd=sd), col = 2, add = T)
  legend('topright', legend = c("Simulated","True density"),
          lwd = c(1, 1), col = c(1,2))
}




```


# 5.3.4 Poisson Process

### (1). Distribution of N(5)

$N(5)\sim Poisson(Z)$, where 

$$Z=\int_0^5\lambda(t)dt=\int_0^5\sqrt{t}+e^{-t}sin(2\pi t)dt$$

$$\int_0^5\sqrt{t}dt=\frac23t^{\frac32}|_0^5=\frac23 5^{\frac32}=7.4536$$

$$
\begin{aligned}
\int_0^5e^{-t}sin(2\pi t)dt
&=-e^{-t}sin(2\pi t)|_0^5+2\pi \int_0^5e^{-t}cos(2\pi t)dt\\
&=0-2\pi \int_0^5cos(2\pi t)de^{-t}\\
&=-(2\pi)^2 \int_0^5 e^{-t}sin(2\pi t)dt-2\pi e^{-t}cos(2\pi t)|_0^5\\
&=-(2\pi)^2 \int_0^5 e^{-t}sin(2\pi t)dt+2\pi(1-e^{-5})\\
so, \int_0^5e^{-t}sin(2\pi t)dt
&=\frac{2\pi(1-e^{-5})}{1+4\pi^4}=0.15418
\end{aligned}
$$
Therefore,

$$Z=\int_0^5\sqrt{t}dt + \int_0^5 e^{-t}sin(2\pi t)dt=7.4536+0.15418=7.60778$$


### (2). Simulation of a nonhomegeneous Poisson process with the thinning method

```{r}
rnhpp <- function(intensity, lammax, tmax) {
  n <- rpois(1, lammax)
  tt <- runif(n, 0, tmax)
  u <- runif(n)
  accept <- u < intensity(tt) / lammax
  sort(tt[accept])
}

lamfun <- function(x) sqrt(x) + exp(-x) * sin(2 * pi * x)

rnhpp(lamfun, 7.6, 5)
```

### (3). Compare the sample density and theoretical density
```{r}
lammax <- 7.6 
tmax <- 5

lamdens <- function(x) lamfun(x) / lammax
simu <-unlist(replicate(1000, rnhpp(lamfun, lammax, tmax)))

hist(simu, freq = F, xlab = "t",
     main = "Simulaiton vs True density of Possion process")
lines(density(simu), col = 1, lwd = 2)
curve(lamdens, 0, 5, add = T, col = 2, lwd = 2)
legend('topleft', legend = c("Simulation","True density"),
       lty = c(1,1), lwd = c(1,1), col=c(1,2),
       cex = 1)

```
